\subsubsection{CommaToDot\label{CommaToDot}\index{CommaToDot}}




\paragraph*{SYNOPSIS\label{CommaToDot_SYNOPSIS}\index{CommaToDot!SYNOPSIS}}


CommaToDot substitutes commas with dots.

\paragraph*{DESCRIPTION\label{CommaToDot_DESCRIPTION}\index{CommaToDot!DESCRIPTION}}


In some countries, e.g. Germany, often the comma is used as the decimal
separator of a number. The method CommaToDot simply changes the comma to
dot to satisfy the database.
CommaToDot() is usually used as a MODIFY-rule in the model file.

\subsubsection{ConvertBool\label{ConvertBool}\index{ConvertBool}}




\paragraph*{SYNOPSIS\label{ConvertBool_SYNOPSIS}\index{ConvertBool!SYNOPSIS}}


\textbf{ConvertBool} converts YyJjNn to the boolean values 1 or 0.

\paragraph*{DESCRIPTION\label{ConvertBool_DESCRIPTION}\index{ConvertBool!DESCRIPTION}}


\textbf{ConvertBool} is usually used as a MODIFY-rule in the model file.

\subsubsection{DotToColon\label{DotToColon}\index{DotToColon}}




\paragraph*{SYNOPSIS\label{DotToColon_SYNOPSIS}\index{DotToColon!SYNOPSIS}}


DotToColon substitutes dots with colons.

\paragraph*{DESCRIPTION\label{DotToColon_DESCRIPTION}\index{DotToColon!DESCRIPTION}}


\textbf{DotToColon} is useful for fast typing of date/time values on numerical keyboards.



Example:
   16.34.00 -$>$  16:34:00



DotToColon() is usually used as a MODIFY-rule in the model file.

\subsubsection{LowerCase\label{LowerCase}\index{LowerCase}}




\paragraph*{SYNOPSIS\label{LowerCase_SYNOPSIS}\index{LowerCase!SYNOPSIS}}


\textbf{LowerCase()} converts the data to lower case

\paragraph*{DESCRIPTION
It is usually used as a MODIFY-rule in the model file.\label{LowerCase_DESCRIPTION_It_is_usually_used_as_a_MODIFY-rule_in_the_model_file_}\index{LowerCase!DESCRIPTION
It is usually used as a MODIFY-rule in the model file.}}
\paragraph*{AUTHORS\label{LowerCase_AUTHORS}\index{LowerCase!AUTHORS}}


Helmut Lichtenberg $<$heli@tzv.fal.de$>$

\subsubsection{SetNow\label{SetNow}\index{SetNow}}




\paragraph*{SYNOPSIS\label{SetNow_SYNOPSIS}\index{SetNow!SYNOPSIS}}


SetNow() returns the current date and time.

\paragraph*{DESCRIPTION\label{SetNow_DESCRIPTION}\index{SetNow!DESCRIPTION}}


SetNow() returns the current date and time. It is usually used as a
MODIFY-rule in the model file. If you need the current time outside the
model file, better use \$apiis-$>$now().

\subsubsection{SetUser\label{SetUser}\index{SetUser}}




\paragraph*{SYNOPSIS\label{SetUser_SYNOPSIS}\index{SetUser!SYNOPSIS}}


SetUser() returns the current user name.

\paragraph*{DESCRIPTION\label{SetUser_DESCRIPTION}\index{SetUser!DESCRIPTION}}


SetUser() sets the passed column extdata to the current user name. It is
usually used as a MODIFY-rule in the model file.

\subsubsection{UpperCase\label{UpperCase}\index{UpperCase}}




\paragraph*{SYNOPSIS\label{UpperCase_SYNOPSIS}\index{UpperCase!SYNOPSIS}}


\textbf{UpperCase()} converts the data to upper case

\paragraph*{DESCRIPTION
It is usually used as a MODIFY-rule in the model file.\label{UpperCase_DESCRIPTION_It_is_usually_used_as_a_MODIFY-rule_in_the_model_file_}\index{UpperCase!DESCRIPTION
It is usually used as a MODIFY-rule in the model file.}}
\paragraph*{AUTHORS\label{UpperCase_AUTHORS}\index{UpperCase!AUTHORS}}


Helmut Lichtenberg $<$heli@tzv.fal.de$>$

